---
title: "Bootstrap"
author: "Sarahy Martinez"
date: "2024-11-19"
output: github_document
---


```{r}
library(tidyverse)
library(modelr)
library(p8105.datasets)


```

## Simulate Data

```{r}
n_samp = 250

sim_df_const = 
  tibble(
    x = rnorm(n_samp, 1, 1),
    error = rnorm(n_samp, 0, 1),
    y = 2 + 3 * x + error
  )

sim_df_nonconst = sim_df_const %>% 
  mutate(
  error = error * .75 * x,
  y = 2 + 3 * x + error
)

```

Plot the data sets

```{r}

sim_df_const %>% 
  ggplot(aes(x =x, y = y))+
  geom_point()+
  geom_smooth(method = "lm")


sim_df_nonconst %>% 
  ggplot(aes(x =x, y = y))+
  geom_point()+
  geom_smooth(method = "lm")


# can fit a linear regression but the key assumptions is fundamentally wrong. Assumptions we make to inference don't match so we will try to solve with bootstrap. Issue comes from figuring out the uncertainty

```

Fit linear regression

```{r}
lm(y~x, data = sim_df_const) %>%  broom::tidy()
lm(y~x, data = sim_df_nonconst) %>%  broom::tidy()

# from our plots we can see that there is alot of uncertainty because of fanning out. We can get estimates and stand deviation if we were to make assumptions but we want to solve the issue of uncertainty by bootstrapping. Issue is that we also don't trust the uncertainty of the estimates.

```



## Draw one bootstrap sample 

```{r}
bootstrap_sample = function(df){   # write a function based on a df
  
  sample_frac(df, replace = TRUE) %>%  # we want the same size bc CI and variance are dependent on the sample size
  
  arrange(x)  # makes easier to look what is going on 
}


```

Check if df works 

```{r}
bootstrap_sample(sim_df_nonconst) %>% 
   ggplot(aes(x =x, y = y))+
  geom_point(alpha = .3)+
  geom_smooth(method = "lm")+
  ylim(-5,16)

# keep running and you'll get different regression lines, analysis is look at the plot but boostrap helps estimate the slope. 


```

Could also 

```{r}

bootstrap_sample(sim_df_nonconst) %>% 
   lm(y ~x, data = .) %>%
  broom::tidy()

# not a cohesive way of analysis
```


## Many samples and analysis 

```{r}

boot_straps =
  tibble(
  strap_number = 1:1000,
  strap_sample = rerun( 1000, bootstrap_sample(sim_df_nonconst))
  )

# once we've gotten here with a dataframe we're right back in place. Can do the same things we did. 


```

Can I run my analysis on these?

```{r}

boot_results =
    boot_straps %>% 
  mutate(
    
    models = map(.x = strap_sample, ~lm( y~x, data = .x)), 
    results = map(models, broom::tidy)
  ) %>% 
  select(strap_number, results) %>% 
  unnest(results)

# code works we have not double checked if they are different but we will trust. next we'll get a collection of results where we will map across models and just run broom tidy on each. 

```


What do I have now?
 Distribution of estimated intercept and slope. So under repeated sampling that the variances are closer to the actual variance if we were to compute the standard error ( standard deviation of estimate) under repeated sampling.
 
 
```{r}

boot_results %>% 
  group_by(term) %>% 
  summarize(
    
    mean_est = mean(estimate),
    sd_est = sd(estimate)
  )


# giving the actual standard error and can we mimic this without making an assumption. Previously we have used a linear model that is a coincidence rather than what happens. If we compare bootstrap this is lower is we assumed constant variance which we have done in this example where we assume constant variance. 

```
 

Constructing a confidence interval and looking at distributions 

```{r}

boot_results %>% 
  filter( term == "x") %>% 
  ggplot(aes(x = estimate))+
  geom_density()
#great bc data of repeated sampling and normally distributed 


```

Construct bootstrap CI

```{r}

boot_results %>% 
  group_by(term) %>% 
  summarize(
    
   ci_lower = quantile(estimate, 0.025),
    ci_upper = quantile(estimate, 0.975)
  )

# we get a CI based on repeated sampling and this gives about 2.91 and 3.31


```

## Bootstrap using modelr
use modelr to increase process of drawing repeated samples. 
Can we simplify anything?

```{r}

sim_df_nonconst %>% 
  bootstrap(1000,id = "strap_number") %>%   #resample and in cross validation we turned into a df, we can do the same thing before 
mutate(   # replaced id with id, we simply just renamed it without having to use the rename function
    
    models = map(.x = strap, ~lm( y~x, data = .x)), 
    results = map(models, broom::tidy)
  ) %>% 
  select(strap_number, results) %>% 
  unnest(results) %>%   
  group_by(term) %>% 
  summarize(
    
    mean_est = mean(estimate),
    sd_est = sd(estimate)
  )

#boostrap says draw me a sample from this df, resample saves memory not new samples. In cross validation we used a df and mcgv but it doesn't work well so we'll do the same thing we've done. 
```

Bootrap works if assumptions are not met but What if your assumptions are met? Does it do something reasonable?

Yes, they both work well

- if assumptions are true we can just use the linear models but we can also do this as well. 

## Revisist NYC Airbnb 

```{r}

data("nyc_airbnb")


nyc_airbnb=
  nyc_airbnb %>% 
  mutate( stars = review_scores_location / 2) %>% 
  rename(
    
    borough = neighbourhood_group,
    neighborhood = neighbourhood) %>% 
  filter(borough != "Staten Island") %>% 
    select(price, stars, borough, neighborhood, room_type)

```


```{r}
# graph first to see if there is any relationship. We see some relationship between stars and price but some non constant variance. We get same estimates but some standard errors that are not correct. 

nyc_airbnb %>% 
  ggplot(aes(x = stars , y = price))+
  geom_point()


```



```{r}

nyc_airbnb %>% 
  filter(borough == "Manhattan") %>% 
  drop_na(stars) %>% 
    ggplot(aes(x = stars , y = price))+
  geom_point()

```

Now we want to bootstrap, take a sample with replacement , fit a regression of price against stars and save the estimated intercept and slope each time. 

```{r}

airbnb_boot_results = nyc_airbnb %>% 
  filter(borough == "Manhattan") %>% 
  drop_na(stars) %>% 
  bootstrap(1000,id = "strap_number") %>%   #generate 1000 samples 
mutate(   
    
    models = map(.x = strap, ~lm( price~stars, data = .x)),  # fit a model to each of these samples price~ stars
    results = map(models, broom::tidy)
  ) %>% 
  select(strap_number, results) %>% 
  unnest(results) 



airbnb_boot_results%>%   
  group_by(term) %>% 
  summarize(
    
    mean_est = mean(estimate),
    sd_est = sd(estimate)
  )



# bootstrap has higher variance bc of the noncontant variance 
```

Compare this to `lm`

```{r}

nyc_airbnb %>% 
  filter(borough == "Manhattan") %>% 
  drop_na(stars) %>% 
  lm(price~stars, data = .) %>% 
  broom::tidy()


```


```{r}


airbnb_boot_results %>% 
  filter( term == "stars") %>% 
  ggplot(aes(x = estimate))+
  geom_density()


# assumig constant variance would think that everything is normally distribute but based on this graph no. We can get uncertainty and distribution of normal sampling wihout having to assume normality with bootstrapping. 

```






































